Modelling transient dynamics of granular slopes: MPM \& DEM

### 2016-10-12 paper.tex
Added Farhang Radjai
MPM friction run-out figures
Fixes changes requested by Revieewer 1
Updated title and added description for low energy regime

\title{Transient dynamics of granular slopes: MPM v DEM}

\author[a]{Krishna Kumar\corref{cor1}}
\author[a,b]{Kenichi Soga}
\author[c]{Jean-Yves Delenne}

\address[a]{Department of Engineering, University of Cambridge, Cambridge CB2
  1PZ, UK}
\address[b]{Department of Civil Engineering, University of California,
  Berkeley, USA}
\address[c]{University of Montpellier II, France}

@@

The DEM sample was composed of $\sim13000$ disks with a uniform distribution of 
diameters by volume fractions ($d_{max} = 1.5 d_{min}$). The mean grain 
diameter and mass are $d\simeq 2.455 $~\si{\mm} and $m\simeq 0.0123$~\si{\kg}, 
respectively. The grains are first poured uniformly into a rectangular box of 
given width and then the right-hand side wall is shifted far to the right 
to allow the grains to spread. A stable granular slope of 13.2~\si{\degree} is 
obtained when all grains come to rest. The initial slope configuration is shown
in~\cref{fig:slope_configuration}. This 
procedure leads to a mean packing fraction $\simeq 0.82$. Soil grains with a 
mean density of 2600~\si{\kg\per\m\cubed} and internal friction coefficient of 
0.4 between grains are considered. For grain-scale simulations, classical DEM
approach was used~\citep{Cundall1979a}.

@@

The flow involves a transient phase with a change in the geometry of the slope 
followed by continuous spreading. The gradient of input energy applied to the 
granular slope mimics a horizontal quake. Despite the creation of a cavity
behind the flowing mass, the granular heap remains in contact with the left
wall irrespective of the input energy.~\Cref{fig:runout_eo_mpm_dem} shows the
normalised run-out distance $(L_f - L_0)/L_0$ and total run-out time $t_f$ as a
function of the input energy $E_0$. Two regimes characterised by a power-law
relation between the run-out distance and time as a function of $E_0$ can be
observed. In the first regime, corresponding to the range of low input energies
$E_0 < 40 \ mgd$, the run-out distance observed varies as $L_f \propto
(E_0)^\alpha$ with $\alpha \simeq 0.206 \pm 0.012$ over nearly one log cycle.
Overall, the run-out distance predicted by the continuum approach matches the
DEM simulations. At very low energies, DEM simulations show longer run-out
distance due to local fluidisation.

@@

\begin{figure}[tbhp]
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/runout_fric_slope}
    \caption{run-out distance}
    \label{fig:runout_fric_slope}
  \end{subfigure}
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/time_fric_slope}
    \caption{duration of run-out}
    \label{fig:time_fric_slope}
  \end{subfigure}
  \caption{MPM simulations of effect of friction on the run-out behaviour of 
  slopes subjected to horizontal excitation.}
  \label{fig:fric_slope}
\end{figure}

>>>>

\title{Modelling transient dynamics of granular slopes: MPM \& DEM}

\author[a]{Krishna Kumar\corref{cor1}}
\author[a,b]{Kenichi Soga}
\author[c]{Jean-Yves Delenne}
\author[c,d]{Farhang Radjai}

\address[a]{Department of Engineering, University of Cambridge, Cambridge CB2
  1PZ, UK}
\address[b]{Department of Civil Engineering, University of California,
  Berkeley, USA}
\address[c]{University of Montpellier II, France}
\address[d]{Massachusetts Institute of Technology, Cambridge, USA}

@@

The DEM sample was composed of $\sim13000$ disks with a uniform distribution of 
diameters by volume fractions ($d_{max} = 1.5 d_{min}$). The mean grain 
diameter and mass are $d\simeq 2.455 $~\si{\mm} and $m\simeq 0.0123$~\si{\kg}, 
respectively. The grains are first poured uniformly into a rectangular box of 
given width and then the right-hand side wall is shifted far to the right 
to allow the grains to spread. A stable granular slope of 13.2~\si{\degree} is 
obtained when all grains come to rest. The initial slope configuration is shown
in~\cref{fig:slope_configuration}. This 
procedure leads to a mean packing fraction (fraction of space filled by disks)
$\simeq 0.82$. Soil grains with a mean density of 2600~\si{\kg\per\m\cubed} and
internal friction coefficient of 0.4 between grains are considered. For 
grain-scale simulations, classical DEM approach was used~\citep{Cundall1979a}.

@@

The flow involves a transient phase with a change in the geometry of the slope 
followed by continuous spreading. The gradient of input energy applied to the 
granular slope mimics a horizontal quake. Despite the creation of a cavity
behind the flowing mass, the granular heap remains in contact with the left
wall irrespective of the input energy.~\Cref{fig:runout_eo_mpm_dem} shows the
normalised run-out distance $(L_f - L_0)/L_0$ and total run-out time $t_f$ as a
function of the input energy $E_0$. Two regimes characterised by a power-law
relation between the run-out distance and time as a function of $E_0$ can be
observed. In the first regime, corresponding to the range of low input energies
$E_0 < 40 \ mgd$, the run-out distance observed varies as $L_f \propto
(E_0)^\alpha$ with $\alpha \simeq 0.206 \pm 0.012$ over nearly one log cycle.
Overall, the run-out distance predicted by the continuum approach matches the
DEM simulations. At very low energies, DEM simulations show longer run-out
distance due to local fluidisation, mainly controlled by failure of the soil
mass at the top of the pile. A critical energy is required, before the entire
soil mass participates in the flow. This results in an almost constant
normalised time in DEM, which is absent in MPM due to the continuum nature.

@@

\begin{figure}[tbhp]
  \centering
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[width=\textwidth]{figs/runout_fric_slope}
    \caption{run-out distance}
    \label{fig:runout_fric_slope}
  \end{subfigure}
  \begin{subfigure}[b]{0.4\textwidth}
    \centering
    \includegraphics[width=0.95\textwidth]{figs/time_fric_slope}
    \caption{duration of run-out}
    \label{fig:time_fric_slope}
  \end{subfigure}
  \caption{MPM simulations of effect of friction on the run-out behaviour of 
  slopes subjected to horizontal excitation.}
  \label{fig:fric_slope}
\end{figure}
